---
title: "NYC Flights"
output: html_notebook
---

This notebook's purpose is to find answers to the week two quiz. It contains basic data analysis of the NYC flights dataset as well as more functions of the R language.

```{r}
# Load the data and the packages
library(statsr)
library(dplyr)
library(ggplot2)

data(nycflights)
names(nycflights)
```


## Exploring the data

```{r}
str(nycflights)
```

```{r}
head(nycflights)
```

## Answering the questions
For this section I am going to create a separate data set with data only from the month of February and only of the flights that headed to San Francisco at that time.

```{r}
# Calculate the amount of flights that fit both categories
sfo_flights <- nycflights %>%
  filter(dest == "SFO" & month == 2)

count(sfo_flights)
```

```{r}
ggplot(data = sfo_flights, aes(x = arr_delay)) +
  geom_histogram(fill="royalblue4", binwidth=5)
```

```{r}
# Finding which carrier has the highest arrival delay IQR
sfo_flights %>%
  group_by(carrier) %>%
  summarise(mean = mean(arr_delay), IQR = IQR(arr_delay), n = n())
```

```{r}
# The month with the highest departure delay (mean, median)
nycflights %>%
  group_by(month) %>%
  summarise(mean = mean(dep_delay), median = median(dep_delay))
```

```{r}
# The carrier with the biggest % of on-time departures
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))

nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate))
```

```{r}
# The aircraft with the highest average speed
nycflights %>%
  group_by(tailnum) %>%
  mutate(avg_speed = distance/(air_time/60)) %>%
  select(tailnum, avg_speed) %>%
  arrange(desc(avg_speed))
```

```{r}
# Scatterplot of the distance against the average speed
ggplot() + geom_point(data = nycflights, aes(x=distance, y=avg_speed, colour=avg_speed), size=1)
```

```{r}
# Create new variable that contains information on whether the arrival was delayed or on time/earlier
nycflights <- nycflights %>%
  mutate(arr_type = ifelse(arr_delay <= 0, "on time", "delayed"))

# Create new variable that explains whether the department was delayed more than 5 minutes, less or equal to 5 minutes or it was on time
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))


delayed_on_time <- nycflights %>% filter(dep_type == "delayed" & arr_type == "on time")
delayed <- nycflights %>% filter(dep_type == "delayed")
perc <- count(delayed_on_time)$n/count(delayed)$n

perc
```

