---
title: "Introduction to R and RStudio"
output: html_notebook
---

This notebook is related to the first course in the Statistics with R specialization and showcases basic functions of the language as well as its most used libraries. I'm going to work with the flights dataset (nycflights13) and perform a quick analysis.

Let's start with exploring the dataset.

```{r}
head(nycflights)
```

```{r}
str(nycflights)
```
```{r}
summary(nycflights)
```
##  Plotting the data

### 1) Simple scatter plot with a gradient
We use geom_point to crate the scatter plot and scale colour gradient to create the gradient. Below we can see arrival delay plotted against the distance.
```{r}
ggplot() + geom_point(data = nycflights, aes(x = distance, y = arr_delay, colour = arr_delay)) + scale_colour_gradient(low="#A9C8F3", high="purple2")
```
### 2) Histogram

```{r}
ggplot(data = nycflights, mapping = aes(x = carrier)) +
  geom_histogram(stat = "count", fill="royalblue4")
```

### 3) Freqpoly
Freqpoly is a line histogram, in this case we display a subset of data: arrival delay lower than 200 grouped by the carrier.
```{r}
ggplot(subset(nycflights, arr_delay < 200), mapping = aes(x = arr_delay)) + 
  geom_freqpoly(mapping = aes(colour = carrier), binwidth = 5)
```
### 4) Box plots
```{r}
ggplot(data = nycflights, mapping = aes(x = distance, y = carrier, fill = carrier)) +
  geom_boxplot()
```

## Creating new features

```{r}
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))

nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate))
```
```{r}
ggplot(data = nycflights, aes(x = origin, fill = dep_type)) +
  geom_bar()
```
## Assignment answers

- 1) Create a new variable called avg_speed and show the fastest planes.
```{r}
nycflights <- nycflights %>% 
  mutate(avg_speed = distance/(air_time * 60))

select(nycflights, avg_speed, tailnum)
```
  
- 2) Make a scatterplot of avg_speed vs. distance. Correct answer: 3, there is a positive association between the variables.

```{r}
ggplot() + geom_point(data = nycflights, aes(x = distance, y = avg_speed, colour = avg_speed)) + scale_colour_gradient(low="#A9C8F3", high="purple2")
```
- 3) Create an on_time variable that depends on the aircraft being on time or earlier than expected.

```{r}
nycflights <- nycflights %>%
  mutate(on_time = ifelse(arr_delay <= 0, "on time", "delayed"))

select(nycflights, on_time, carrier)
```